Tulai

### 2024-10-23 paper/paper.qmd
CA analysis added

@@

>>>>

```{r}
# Set the environment variable for GDAL
Sys.setenv(GDAL_CONFIG = "/path/to/your/conda/env/bin/gdal-config")

# Install the 'terra' package
install.packages("terra")

@@

# Correspondence Analysis (CA)
```{r}
# Example of a contingency table
lithic_data <- matrix(c(30, 20, 50, 
                        15, 25, 35,
                        45, 10, 30), 
                      nrow = 3, byrow = TRUE)
colnames(lithic_data) <- c("Cores", "Blades", "Flakes")
rownames(lithic_data) <- c("Unit1", "Unit2", "Unit3")

```{r}
# Perform Correspondence Analysis
res.ca <- CA(lithic_data, graph = FALSE)

# View the summary results
summary(res.ca)

```{r}
# Plot the CA biplot
fviz_ca_biplot(res.ca, repel = TRUE)

# contribution plots
```{r}
# Example lithic data matrix
lithic_data <- matrix(c(30, 20, 50, 
                        15, 25, 35,
                        45, 10, 30), 
                      nrow = 3, byrow = TRUE)
colnames(lithic_data) <- c("Cores", "Blades", "Flakes")
rownames(lithic_data) <- c("Unit1", "Unit2", "Unit3")

# Perform Correspondence Analysis (CA)
res.ca <- CA(lithic_data, graph = FALSE)

```{r}
# Contribution of rows to Dimension 1
fviz_contrib(res.ca, choice = "row", axes = 1, top = 3)

# Contribution of rows to Dimension 2
fviz_contrib(res.ca, choice = "row", axes = 2, top = 3)
```

```{r}
# Contribution of columns to Dimension 1
fviz_contrib(res.ca, choice = "col", axes = 1, top = 3)

# Contribution of columns to Dimension 2
fviz_contrib(res.ca, choice = "col", axes = 2, top = 3)

```{r}
# Combine the top contributing rows for both Dim1 and Dim2
fviz_contrib(res.ca, choice = "row", axes = 1:2, top = 3)

# Combine the top contributing columns for both Dim1 and Dim2
fviz_contrib(res.ca, choice = "col", axes = 1:2, top = 3)

```{r}
# Save the contribution plot for rows
ggsave("contribution_plot_rows_Dim1.png", plot = fviz_contrib(res.ca, choice = "row", axes = 1, top = 3))

# Save the contribution plot for columns
ggsave("contribution_plot_cols_Dim1.png", plot = fviz_contrib(res.ca, choice = "col", axes = 1, top = 3))
